---
title: "p8105_hw5_rr3423.Rmd"
author: "Ragyie Rawal"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## PROBLEM 1 

```{r import_data}
homicide_data = read_csv("data/homicide-data.csv", na = c("", "Unknown"))
```

### Describing raw data 

```{r describing_data}
homicide_variable = 
  homicide_data %>% 
  names()

homicide_columns = 
  homicide_data %>% 
  ncol()

homicide_rows = 
  homicide_data %>% 
  nrow()
```

The **key variables** in the raw homicide dataset are `r homicide_variable`. The **number of columns** in the raw homicide dataset are `r homicide_columns`. The **number of rows** in the raw homicide dataset are `r homicide_rows`. There is data provided for a total of **`r homicide_rows` homicides** from 50 large U.S. cities. The raw dataset contained "Unknown" and blank data values, which I replaced with "NA" values when importing the csv file. 

### Creating city-state and resolution variables

```{r tidy_data}
homicide_df = 
  homicide_data %>%
  mutate(
    city_state = str_c(city, state),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved"
    )) %>% 
  relocate(city_state) %>% 
  filter(city_state != "TulsaAL")
```

### Summary of the total number of homicides and the number of unsolved homicides within cities

```{r summarize_cities, message = FALSE, warning = FALSE}
# summarizing total number of homicides within cities 
homicide_df %>% 
  group_by(city_state) %>% 
  summarize(
    unsolved = sum(resolution == "unsolved"), 
    total = n()
  ) %>% 
  knitr::kable(
    caption = "**Table 1: Total number of homicides and unsolved homicides within cities**")
```

### Focusing on Baltimore, MD 

```{r baltimore_analysis}
baltimore_df = 
  homicide_df %>% 
  filter(city_state == "BaltimoreMD")

baltimore_summary = 
  baltimore_df %>% 
  summarize(
    unsolved = sum(resolution == "unsolved"),
    n = n()
  )

# prop.test to estimate proportion of unsolved homicides in baltimore 
baltimore_test = 
  prop.test(
  x = baltimore_summary %>% pull(unsolved),
  n = baltimore_summary %>% pull(n)
)

# applying broom::tidy
baltimore_test %>% 
  broom::tidy()
```

From the resulting tidy dataframe, the estimated proportion is 0.646 and the 95% confidence interval is (0.628, 0.663). This means that approximately 64.6% of homicides in Baltimore are unsolved. 

### Creating prop.test function 

```{r prop_test_function}
# creating a function 
prop_test_function = function(city_df) {
  
  city_summary = 
    city_df %>% 
    summarize(
      unsolved = sum(resolution == "unsolved"),
      n = n()
    )

  city_test = 
   prop.test(
    x = city_summary %>% pull(unsolved),
    n = city_summary %>% pull(n)
  )
  
  return(city_test)
  
}

# testing function on some cities
prop_test_function(baltimore_df)

homicide_df %>% 
  filter(city_state == "AlbuquerqueNM") %>% 
  prop_test_function
```

### Iterating across cities 

```{r iterating_cities}
# tidy dataframe with estimated proportions and CIs for each city 
results_df = 
  homicide_df %>% 
  nest(data = uid:resolution) %>%
  mutate(
    test_results = map(data, prop_test_function),
    tidy_results = map(test_results, broom::tidy)
  ) %>% 
  select(city_state, tidy_results) %>% 
  unnest(tidy_results) %>% 
  select(city_state, estimate, starts_with("conf"))
```

### Plot showing estimates and CIs for each city 

```{r plot}
results_df %>% 
  mutate(city_state = fct_reorder(city_state, estimate)) %>% 
  ggplot(aes(x = city_state, y = estimate)) +
  geom_point() +
  geom_errorbar(aes(ymin = conf.low, ymax = conf.high)) +
  theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust = 1)) + 
  labs(
    title = "Estimates and confidence intervals for each city",
    x = "City", 
    y = "Estimate"
  )
```


## PROBLEM 2 


## PROBLEM 3 
