---
title: "p8105_hw5_rr3423.Rmd"
author: "Ragyie Rawal"
output: github_document
---

```{r setup, include = FALSE}
library(tidyverse)
library(p8105.datasets)
library(viridis)

knitr::opts_chunk$set(
  fig.width = 6,
  fig.asp = .6,
  out.width = "90%"
)

theme_set(theme_minimal() + theme(legend.position = "bottom"))

options(
  ggplot2.continuous.colour = "viridis",
  ggplot2.continuous.fill = "viridis"
)

scale_colour_discrete = scale_colour_viridis_d
scale_fill_discrete = scale_fill_viridis_d
```


## PROBLEM 1 

```{r import_data}
# importing homicide dataset 
homicide_data = read_csv("data/homicide-data.csv", na = c("", "Unknown"))
```

### Describing raw data 

```{r describing_data}
homicide_variable = 
  homicide_data %>% 
  names()

homicide_columns = 
  homicide_data %>% 
  ncol()

homicide_rows = 
  homicide_data %>% 
  nrow()
```

The **key variables** in the raw homicide dataset are `r homicide_variable`. The **number of columns** in the raw homicide dataset are `r homicide_columns`. The **number of rows** in the raw homicide dataset are `r homicide_rows`. There is data provided for a total of **`r homicide_rows` homicides** from 50 large U.S. cities. The raw dataset contained "Unknown" and blank data values, which I replaced with "NA" values when importing the csv file. 

### Tidying data, creating city-state variable

```{r tidy_data}
homicide_df = 
  homicide_data %>%
  mutate(
    city_state = str_c(city, state),
    resolution = case_when(
      disposition == "Closed without arrest" ~ "unsolved",
      disposition == "Open/No arrest" ~ "unsolved",
      disposition == "Closed by arrest" ~ "solved"
    )) %>% 
  relocate(city_state)
```

